New lower bounds for Schur and weak Schur numbers

### 2021-06-13 article/introduction.tex
modifs
valeurs exactes en couleur

\qquad We are interested in partioning the set of integers \(\{1, ..., p\}\) into \(n\) subsets such that there is no 
subset containing three integers \(x\), \(y\) and \(z\) verifying \(x + y = z\). We say these subsets are 
\textit{sum-free}. If we add the hypothesis \(x \neq y\), we say the subsets are \textit{weakly sum-free}. The 
greatest \(p\) for which there exists a partition into \(n\) sum-free subsets is called the \(n^{\text{th}}\) Schur 
number and is denoted \(S(n)\) \cite{Schur1917}. Likewise for weakly sum-free partitions we define \(\WS(n)\) 
the \(n^{\text{th}}\) weak Schur number \cite{Irving1973}. Only the first values of these sequences are known. 
In this study, we improve the lower bounds on these numbers.

@@

\setlength{\tabcolsep}{4pt}

\begin{table}[H]
  \centering
\begin{tabular}{|*{13}{c|}}
    \hline
    \(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
    \hline
    \multirow{2}{*}{Before Rowley} & 1* & 4* & 13* & 44* & 160* & 536 & 1\,680 & 5\,041 & 15\,124 & 51\,120 & 172\,216 & 575\,664 \\
      & & & & & \cite{Heule2017} & \cite{Fredricksen} & \cite{Fredricksen} & \cite{AbbottHanson} & 
    \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
    \hline
    \begin{tabular}{@{}c@{}}Rowley \\ \cite{RowleyRamsey}\end{tabular} & & & & & & & & 5\,286 & 17\,694 & 
    60\,320 & 201\,696 & 631\,840 \\
    \hline
    \hyperref[Schur]{\textbf{Our results}}  & & & & & & & & & \textbf{17\,803} & \textbf{60\,948} & \textbf{203\,828} & \textbf{638\,548} \\
    \hline
\end{tabular}
\caption{Comparison of lower bounds for Schur numbers}
\label{table:1}
\end{table}
\vspace{2.5ex}

\begin{table}[H]
  \centering
\begin{tabular}{|*{13}{c|}}
    \hline
    \(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
    \hline
    \multirow{2}{*}{Before Rowley} & 2* & 8* & 23* & 66* & 196 & 582 & 1\,740 & 5\,201 & 15\,596 & 51\,520 & 172\,216 & 575\,664 \\
      & & & & & \cite{ELIAHOU2012175} &\cite{EliahouBook} & \cite{Rafilipojaona} & \cite{Rafilipojaona} & 
    \cite{Rafilipojaona} & \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
    \hline
    \begin{tabular}{@{}c@{}}Rowley \\ \cite{RowleyWS}\end{tabular} & & & & & & 642 & 2\,146 & 6\,976 & 21\,848 
    & 70\,778 & 241\,282 & 806\,786 \\
    \hline
    \hyperref[WeakSchur]{\textbf{Our results}} & & & & & & & & & \textbf{22\,536} & \textbf{71\,214} & \textbf{243\,794} & \textbf{815\,314} \\
    \hline
\end{tabular}
\caption{Comparison of lower bounds for weak Schur numbers}
\label{table:2}
\end{table}

\begin{center}
* denotes the exact value
\end{center}

@@

\qquad We start by defining sum-free and weakly sum-free subsets to introduce regular and weak Schur numbers.

>>>>

\qquad We are interested in partioning the set of integers \(\{1, ..., p\}\) into \(n\) subsets such that there is no 
subset containing three integers \(x\), \(y\) and \(z\) verifying \(x + y = z\). We say these subsets are 
\textit{sum-free}. If we add the hypothesis \(x \neq y\), we say the subsets are \textit{weakly sum-free}. The 
greatest \(p\) for which there exists a partition into \(n\) sum-free subsets is called the \(n^{\text{th}}\) Schur 
number and is denoted \(S(n)\) \cite{Schur1917}. Likewise for weakly sum-free partitions we define \(\WS(n)\) 
the \(n^{\text{th}}\) weak Schur number \cite{Irving1973}. Values of \(S(n)\) and \(\WS(n)\) are known for small \(n\) only.

@@

\par The hereunder tables recap the lower bounds respectively for Schur and weak Schur numbers. Gray cells 
indicate the exact values.

\begin{table}[H]
  \centering
\begin{NiceTabular}{*{13}{c}}[vlines]
\CodeBefore
	\cellcolor{gray!40}{2-2}
	\cellcolor{gray!40}{2-3}
	\cellcolor{gray!40}{2-4}
	\cellcolor{gray!40}{2-5}
	\cellcolor{gray!40}{2-6}
	\cellcolor{gray!40}{3-2}
	\cellcolor{gray!40}{3-3}
	\cellcolor{gray!40}{3-4}
	\cellcolor{gray!40}{3-5}
	\cellcolor{gray!40}{3-6}
\Body
	\hline
	\(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
	\hline
	\Block{2-1}{Before Rowley} & 1 & 4 & 13 & 44 & 160 & 536 & 1\,680 & 5\,041 & 15\,124 & 51\,120 & 172\,216 & 575\,664 \\
	& & & & & \cite{Heule2017} & \cite{Fredricksen} & \cite{Fredricksen} & \cite{AbbottHanson} & \cite{AbbottHanson} & 
		\cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
	\hline
	Rowley & \Block{2-1}{} & \Block{2-1}{} & \Block{2-1}{} & \Block{2-1}{} & \Block{2-1}{} & \Block{2-1}{} & \Block{2-1}{} & 
		\Block{2-1}{5\,286} & \Block{2-1}{17\,694} & \Block{2-1}{60\,320} & \Block{2-1}{201\,696} & \Block{2-1}{631\,840} \\
	\cite{RowleyRamsey} & & & & & & & & & & & & \\
	\hline
	\hyperref[Schur]{\textbf{Our results}}  & & & & & & & & & \textbf{17\,803} & \textbf{60\,948} & \textbf{203\,828} & \textbf{638\,548} \\
	\hline
\end{NiceTabular}
\caption{Comparison of lower bounds for Schur numbers}
\label{table:1}
\end{table}

\begin{table}[H]
  \centering
\begin{NiceTabular}{*{13}{c}}[vlines]
\CodeBefore
	\cellcolor{gray!40}{2-2}
	\cellcolor{gray!40}{2-3}
	\cellcolor{gray!40}{2-4}
	\cellcolor{gray!40}{2-5}
	\cellcolor{gray!40}{3-2}
	\cellcolor{gray!40}{3-3}
	\cellcolor{gray!40}{3-4}
	\cellcolor{gray!40}{3-5}
\Body
	\hline
	\(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
	\hline
	\Block{2-1}{Before Rowley} & 2 & 8 & 23 & 66 & 196 & 582 & 1\,740 & 5\,201 & 15\,596 & 51\,520 & 172\,216 & 575\,664 \\
	& & & & & \cite{ELIAHOU2012175} &\cite{EliahouBook} & \cite{Rafilipojaona} & \cite{Rafilipojaona} & \cite{Rafilipojaona} & 
		\cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
	\hline
	Rowley & \Block{2-1}{} & \Block{2-1}{} & \Block{2-1}{} & \Block{2-1}{} & \Block{2-1}{} & \Block{2-1}{642} & \Block{2-1}{2\,146} & 
		\Block{2-1}{6\,976} & \Block{2-1}{21\,848} & \Block{2-1}{70\,778} & \Block{2-1}{241\,282} & \Block{2-1}{806\,786} \\
	\cite{RowleyWS} & & & & & & & & & & & & \\
	\hline
	\hyperref[Schur]{\textbf{Our results}}  & & & & & & & & & \textbf{22\,536} & \textbf{71\,214} & \textbf{243\,794} & \textbf{815\,314} \\
	\hline
\end{NiceTabular}
\caption{Comparison of lower bounds for weak Schur numbers}
\label{table:2}
\end{table}

@@

\qquad We start by defining sum-free and weakly sum-free subsets to introduce regular and weak Schur numbers. 
We denote \(\mathbb{N} = \{0, 1, 2, ...\}\) and \(\mathbb{N}^* =\mathbb{N} \backslash \{0\}\).

### 2021-06-13 article/weakschur.tex
valeurs exactes en couleur

\qquad In this section, we first gave a new construction which can be seen as an equivalent for weakly sum-free partitions of Abott
and Hanson's construction for sum-free partitions. We then  introduced WS-templates and generalized this construction. This
allows us to find new lower bounds and new inequalities for weak Schur numbers. One may notice the significant gap
between the former lower bounds for weak Schur numbers obtained by conducting a computer search and the new lower bounds
obtained with WS-templates.

>>>>

\qquad We started by giving a new construction which can be seen as an equivalent for weakly sum-free partitions of Abott
and Hanson's construction for sum-free partitions. We then generalized this construction by introducing WS-templates. This
allows us to find new lower bounds and new inequalities for weak Schur numbers. One may notice the significant gap
between the former lower bounds for weak Schur numbers obtained by conducting a computer search and the new lower bounds
obtained with WS-templates.
